---
title: "MSCquartets"
output: html_notebook
---

This is an exploration of [MSCquartets](https://cran.r-project.org/web/packages/MSCquartets/MSCquartets.pdf).

The goal of this R notebook is to:
1. determine what functions given by the MSCquartets package are relevant
2. how to interpret them
3. determine their input/output

startup: import libraries
```{r}
library(MSCquartets)
library(ape)
```

data input: collection of gene trees

Computing Quartet Concordance Factors:
1. take set of gene trees, and for sets of 4 taxa abcd:
  a. possibilities broken down into 
  ab|cd, ac|bd, ad|bc, abcd   (unresolved/star)
2. unresolved/stars have prob 0 under MSC, can be thrown out or used to add n_abcd / 3 to all other counts

Step 1: import trees data using '''ape''' package's [read.tree](https://www.rdocumentation.org/packages/ape/versions/5.4-1/topics/read.tree) function.

```{r}
setwd("~/Github/phylo-microbes/data/")
fileName = "knownGT/n10_n100/1_astral.in" 
gene_trees = read.tree(file = fileName)
```

Step 2: form a quartet table using MSCquartets '''quartetTable'''. 

The resulting quartet table can be resolved (scrubbing unresolved/star trees) with '''resolveQuartetTable'''.
```{r}
QT = quartetTable(gene_trees, taxonnames = NULL, epsilon = 0, random = 0)

QTR = quartetTableResolved(QT, omit = F)

```

hybridization / gene duplication/loss would result in probabilities diverging from the lines connecting the vertices to the centroid of the simplex plot.

Step 3: use resolved quartet table to produce p-vals under different hypotheses. 

'''quartetTreeTestInd''' - gene trees fit MSC model
model = "T3", (T1 requires species tree topology). This outputs p-values for each quartet.

Output can be used to plot the location on a simplex, and mark as reject/fail to reject based on p-values.

```{r}
MSC = quartetTreeTestInd(QTR, model = "T3")
quartetTestPlot(MSC, test = "T3", alpha = 0.05, beta = 1, cex = 1)
```

Quartet concordance factors can also be visualized in simplex form sans p-test (not as valuable)

```{r}
simplexPrepare("T3",fileName)

num_rows = nrow(QTR)
for (row in 1:num_rows) {
  point = c(QTR[row,"12|34"], QTR[row,"13|24"], QTR[row,"14|23"])
  simplexPoint(point,pch=3,col="blue")
}
```




